---
title: "Edge-Score"
author: "Zhesi"
date: "2026-09-18"
output: html_document
---

This is the tutorial to implement Edge-Score enrichment analysis method.\
Before running, please set up your working directory.

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/scratch/project_2011179/code/github/")  # Please change your woking directory here.
```

Note: in .R script, please set your working directory by `setwd("[your directory]")`.

You can run Edge-Score enrichment analysis in three ways:\
**A.** run through integrated miREA() function.\
**B.** run by generating ranked MGI list through miRNA, gene differential expression analysis result and miRNA, gene expression matrix with common samples.\
**C.** run by providing self-defined ranked MGI list.\
In the following tutorial, we will go through each accordingly.

**Required input:**\
**1. background_MGI:** A two-column data frame with column names miRNA and gene.\
**2. pathway:** A two-column data frame with column names pathway and gene.\
**3. mir_DEdata (required for A, B):** A four-column data frame containing miRNA, log2FoldChange, stat, and padj.\
**4. gene_DEdata (required for A, B):** A four-column data frame containing gene, log2FoldChange, stat, and padj.\
**5. mir_mat (required for A, B):** A raw miRNA expression matrix, with each row representing a miRNA and each column representing a sample.\
**6. gene_mat (required for A, B):** A raw gene expression matrix, with each row representing a gene and each column representing a sample.\
**7. MGIList (required for C):** A named vector of MGIs ranked by certain statistics. The MGIs should in the format of [miRNA]:[gene].

### A. run through miREA().

```{r}
load("function.RData") # it contains all functions included in the R/.
source("R/lib.R")
methods <- "Edge_Score"

mir_DEdata <- read.csv("data/raw_data/cancer_data/DEmiR/BLCA_DEmiR.csv")
mir_DEdata <- mir_DEdata %>% select(miRNA, log2FoldChange, stat, padj)
gene_DEdata <- read.csv("data/raw_data/cancer_data/DEG/BLCA_DEG.csv")
gene_DEdata <- gene_DEdata %>% select(gene, log2FoldChange, stat, padj)
gene_mat <- read.csv("data/raw_data/cancer_data/paired/BLCA_gene_TP.csv", header = TRUE, check.names = FALSE)
mir_mat <- read.csv("data/raw_data/cancer_data/paired/BLCA_mir_TP.csv", header = TRUE, check.names = FALSE)

path_name <-  "hallmark"  # or "Reactome"
pathway <- read.csv(paste0("data/raw_data/pathway/", path_name, "/", path_name, "_gene.csv"), header = TRUE)

input_data <- get_all_input_data(methods = methods, pathway = pathway, mir_DEdata = mir_DEdata, gene_DEdata = gene_DEdata, gene_mat = gene_mat, mir_mat = mir_mat)
result <- miREA(methods = methods, input_data = input_data)
```

It will generate a list named result, which contains the enrichment analysis result (`result$result$Edge_Score`), calculation time (`result$time`), methods (`result$valid_methods`), and parameters (`result$parameter`).

### B. run by generating ranked MGI list through DE and expression data.
```{r}
# load functions
source("R/Enrichment_Edge-Score.R")
source("R/get_input_data.R")  # calculate MGI strength score
source("R/global_variable.R")
source("R/lib.R")

# load data
mir_DEdata <- read.csv("data/raw_data/cancer_data/DEmiR/BLCA_DEmiR.csv")
gene_DEdata <- read.csv("data/raw_data/cancer_data/DEG/BLCA_DEG.csv")
gene_mat <- read.csv("data/raw_data/cancer_data/paired/BLCA_gene_TP.csv", header = TRUE, check.names = FALSE)
mir_mat <- read.csv("data/raw_data/cancer_data/paired/BLCA_mir_TP.csv", header = TRUE, check.names = FALSE)

path_name <-  "hallmark"  # or "Reactome"
pathway <- read.csv(paste0("data/raw_data/pathway/", path_name, "/", path_name, "_gene.csv"), header = TRUE)
background_MGI <- read.csv("data/raw_data/background/background_MGI.csv")

mir_DEdata <- data.table::as.data.table(mir_DEdata)
mir_DEdata <- mir_DEdata[, .(miRNA, log2FoldChange, stat, padj)]
colnames(mir_DEdata) <- c("miRNA", "mirLogFC", "mirStat", "mirPadj")
gene_DEdata <- data.table::as.data.table(gene_DEdata)
gene_DEdata <- gene_DEdata[, .(gene, log2FoldChange, stat, padj)]
colnames(gene_DEdata) <- c("gene", "geneLogFC", "geneStat", "genePadj")

background_MGI <- data.table::as.data.table(background_MGI)
pathway <- data.table::as.data.table(pathway)

## build pathway MGI list
pathway_genes <- pathway[, .(genes = list(gene)), by = pathway]
pathway_list <- lapply(seq_len(nrow(pathway_genes)), function(i) {
  pw <- pathway_genes$pathway[i]
  genes <- pathway_genes$genes[[i]]
  background_MGI[gene %in% genes, .(pathway = pw, miRNA, gene)]
})
pathway_MGI <- unique(rbindlist(pathway_list))

## build MGI dict and calculate MGI strength score
dict <- merge(background_MGI, mir_DEdata[, .(miRNA, mirLogFC, mirPadj)], by = "miRNA", all = FALSE)
dict <- merge(dict, gene_DEdata[, .(gene, geneLogFC, genePadj)], by = "gene", all = FALSE)

cor <- data.table::as.data.table(calc_MGI_corr(dict = dict, mir_mat = mir_mat, gene_mat = gene_mat))
dict <- merge(dict, cor, by = c("miRNA", "gene"), all = FALSE)
dict <- na.omit(dict)
data.table::setcolorder(dict, c("miRNA", "gene", "cor", "corPadj", "mirLogFC", "mirPadj", "geneLogFC", "genePadj"))
dict[, ratio := (geneLogFC / mirLogFC) * ((1-genePadj)/(1+mirPadj))]
dict[, normratio := ranknorm(ratio)]
dict[, `:=`(
  strength = cor * normratio,
  MGI = paste0(miRNA, ":", gene)
)]
setorder(dict, strength)
dict <- unique(dict, by = "MGI")

MGIList <- setNames(as.numeric(dict$strength), dict$MGI)

## enrichment analysis
result <- Edge_Score(MGIList = MGIList, pathway = pathway_MGI, pvalueCutoff = 1)
```

### C. run by providing self-defined interested MGI list.
```{r}
# load functions
source("R/Enrichment_Edge-Score.R")
source("R/global_variable.R")
source("R/lib.R")

# load data
path_name <-  "hallmark"  # or "Reactome"
pathway <- read.csv(paste0("data/raw_data/pathway/", path_name, "/", path_name, "_gene.csv"), header = TRUE)
background_MGI <- read.csv("data/raw_data/background/background_MGI.csv")

background_MGI <- data.table::as.data.table(background_MGI)
pathway <- data.table::as.data.table(pathway)

## build pathway MGI list
pathway_genes <- pathway[, .(genes = list(gene)), by = pathway]
pathway_list <- lapply(seq_len(nrow(pathway_genes)), function(i) {
  pw <- pathway_genes$pathway[i]
  genes <- pathway_genes$genes[[i]]
  background_MGI[gene %in% genes, .(pathway = pw, miRNA, gene)]
})
pathway_MGI <- unique(rbindlist(pathway_list))

## specify DEMGI list
DEMGI <- c("hsa-let-7c-3p:CASZ1", "hsa-let-7c-3p:E2F3", "hsa-let-7c-3p:PTPRS", "hsa-let-7c-5p:ASB16", "hsa-let-7c-5p:DIABLO", "hsa-let-7c-5p:GMNN", "hsa-miR-1-3p:BCAP29", "hsa-miR-1-3p:CENPL")
MGIList <- setNames(length(DEMGI):1, DEMGI)

## enrichment analysis
result <- Edge_Score(MGIList = MGIList, pathway = pathway_MGI, pvalueCutoff = 1)
```
