---
title: "MiR-Score"
author: "Zhesi"
date: "2026-09-18"
output: html_document
---

This is the tutorial to implement MiR-Score enrichment analysis method.\
Before running, please set up your working directory.

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/scratch/project_2011179/code/github/")  # Please change your woking directory here.
```

Note: in .R script, please set your working directory by `setwd("[your directory]")`.

You can run MiR-Score enrichment analysis in three ways:\
**A.** run through integrated miREA() function.\
**B.** run by providing ranked miRNA list through miRNA differential expression analysis result.\
**C.** run by providing self-defined ranked miRNA list.\
In the following tutorial, we will go through each accordingly.

**Required input:**\
**1. background_MGI:** A two-column data frame with column names miRNA and gene.\
**2. pathway:** A two-column data frame with column names pathway and gene.\
**3. mir_DEdata (required for A, B):** A four-column data frame containing miRNA, log2FoldChange, stat, and padj.\
**4. miRList (required for C):** A named vector of miRNAs ranked by certain statistics.

### A. run through miREA().

```{r}
load("function.RData") # it contains all functions included in the R/.
source("R/lib.R")
methods <- "MiR_Score"

mir_DEdata <- read.csv("data/raw_data/cancer_data/DEmiR/BLCA_DEmiR.csv")
mir_DEdata <- mir_DEdata %>% select(miRNA, log2FoldChange, stat, padj)


path_name <- "Reactome" # or "hallmark"
pathway <- read.csv(paste0("data/raw_data/pathway/", path_name, "/", path_name, "_gene.csv"), header = TRUE)

input_data <- get_all_input_data(methods = methods, pathway = pathway, mir_DEdata = mir_DEdata)
result <- miREA(methods = methods, input_data = input_data)
```

It will generate a list named result, which contains the enrichment analysis result (`result$result$MiR_Score`), calculation time (`result$time`), methods (`result$valid_methods`), and parameters (`result$parameter`).

### B. run by providing miRNA differential expression analysis result.

Load miRNA differential expression analysis result to generate ranked gene list.

```{r}
# load functions
source("R/Enrichment_miRNA_based.R")
source("R/global_variable.R")
source("R/lib.R")

# load MGI background and pathway gene set
background_MGI <- read.csv("data/raw_data/background/background_MGI.csv")
path_name <- "Reactome" # or "hallmark"
pathway <- read.csv(paste0("data/raw_data/pathway/", path_name, "/", path_name, "_gene.csv"), header = TRUE)

background_MGI <- data.table::as.data.table(background_MGI)
pathway <- data.table::as.data.table(pathway)

# build input miRNA list
mir_DEdata <- read.csv("data/raw_data/cancer_data/DEmiR/BLCA_DEmiR.csv")
mir_DEdata <- data.table::as.data.table(mir_DEdata)
miRList <- mir_DEdata[, setNames(stat, miRNA)]
miRList <- sort(miRList, decreasing = TRUE)

# transform pathway gene set to pathway miRNA set
pathway_genes <- pathway[, .(genes = list(gene)), by = pathway]
pathway_list <- lapply(seq_len(nrow(pathway_genes)), function(i) {
  pw <- pathway_genes$pathway[i]
  genes <- pathway_genes$genes[[i]]
  background_MGI[gene %in% genes, .(pathway = pw, miRNA, gene)]
})
pathway_MGI <- unique(rbindlist(pathway_list))
pathway_miR <- unique(pathway_MGI[, .(pathway, miRNA)])

## enrichment analysis
result <- MiR_Score(miRList = miRList, pathway = pathway_miR, pvalueCutoff = 1)
```

### C. run by providing self-defined ranked miRNA list.

You can also specify a ranked miRNA list by yourself.

```{r}
# load functions
source("R/Enrichment_miRNA_based.R")
source("R/global_variable.R")
source("R/lib.R")

# load MGI background and pathway gene set
background_MGI <- read.csv("data/raw_data/background/background_MGI.csv")
path_name <- "Reactome" # or "hallmark"
pathway <- read.csv(paste0("data/raw_data/pathway/", path_name, "/", path_name, "_gene.csv"), header = TRUE)

background_MGI <- data.table::as.data.table(background_MGI)
pathway <- data.table::as.data.table(pathway)

# provide ranked miRNA list
miRs <- c("hsa-miR-210-3p", "hsa-miR-139-3p", "hsa-miR-96-5p", "hsa-miR-195-5p", "hsa-miR-30a-3p", "hsa-miR-28-3p", "hsa-miR-18a-5p", "hsa-miR-143-3p", "hsa-miR-1307-5p", "hsa-miR-17-5p")
miRList <- setNames(length(miRs):1, miRs)

# transform pathway gene set to pathway miRNA set
pathway_genes <- pathway[, .(genes = list(gene)), by = pathway]
pathway_list <- lapply(seq_len(nrow(pathway_genes)), function(i) {
  pw <- pathway_genes$pathway[i]
  genes <- pathway_genes$genes[[i]]
  background_MGI[gene %in% genes, .(pathway = pw, miRNA, gene)]
})
pathway_MGI <- unique(rbindlist(pathway_list))
pathway_miR <- unique(pathway_MGI[, .(pathway, miRNA)])

## enrichment analysis
result <- MiR_Score(miRList = miRList, pathway = pathway_miR, pvalueCutoff = 1)
```