---
title: "TG-ORA"
author: "Zhesi"
date: "2026-09-18"
output: html_document
---

This is the tutorial to implement TG-ORA enrichment analysis method.\
Before running, please set up your working directory

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/scratch/project_2011179/code/github/")  # Please change your woking directory here.
```

Note: in .R script, please set your working directory by `setwd("[your directory]")`.

You can run TG-ORA enrichment analysis in three ways:\
**A.** run through integrated miREA() function.\
**B.** run by providing input gene list through miRNA differential expression analysis result.\
**C.** run by specifying an interested miRNA list.\
In the following tutorial, we will go through each accordingly.

**Required input:**\
**1. background_MGI:** A two-column data frame with column names miRNA and gene.\
**2. pathway:** A two-column data frame with column names pathway and gene.\
**3. mir_DEdata (required in A, B):** A four-column data frame containing miRNA, log2FoldChange, stat, and padj.\
**4. miR (required in C):** a character vector containing miRNAs of interest.

### A. run through miREA().

```{r}
load("function.RData") # it contains all functions included in the R/.
source("R/lib.R")
methods <- "TG_ORA"

mir_DEdata <- read.csv("data/raw_data/cancer_data/DEmiR/BLCA_DEmiR.csv")
mir_DEdata <- mir_DEdata %>% select(miRNA, log2FoldChange, stat, padj)

path_name <- "Reactome" # or "hallmark"
pathway <- read.csv(paste0("data/raw_data/pathway/", path_name, "/", path_name, "_gene.csv"), header = TRUE)

input_data <- get_all_input_data(methods = methods, pathway = pathway, mir_DEdata = mir_DEdata)
result <- miREA(methods = methods, input_data = input_data)
```

It will generate a list named result, which contains the enrichment analysis result (`result$result$TG_ORA`), calculation time (`result$time`), methods (`result$valid_methods`), and parameters (`result$parameter`).

### B. run by providing miRNA differential expression analysis result.

Load miRNA differential expression analysis result, identify DEmiRs, and define input gene list as DEmiRs' target genes.

```{r}
# load functions
source("R/Enrichment_TG_based.R")
source("R/global_variable.R")
source("R/lib.R")

# load MGI background and pathway gene set
background_MGI <- read.csv("data/raw_data/background/background_MGI.csv")
path_name <- "Reactome" # or "hallmark"
pathway <- read.csv(paste0("data/raw_data/pathway/", path_name, "/", path_name, "_gene.csv"), header = TRUE)


background_MGI <- data.table::as.data.table(background_MGI)
pathway <- data.table::as.data.table(pathway)

# generate TG from miRNA DE data
mir_DEdata <- read.csv("data/raw_data/cancer_data/DEmiR/BLCA_DEmiR.csv")
mir_DEdata <- data.table::as.data.table(mir_DEdata)

DEmiRs <- mir_DEdata[abs(log2FoldChange) > 1 & padj < 0.05, unique(miRNA)]
TG <- background_MGI[miRNA %in% DEmiRs, unique(gene)]

# run TG-ORA
result <- TG_ORA(TG = TG, pathway = pathway, pvalueCutoff = 1)
```

### C. run by specifying an interested miRNA list.

if you have a list of interested miRNA directly, you could transform interested miRNAs to their target genes as input gene list.

```{r}
# load functions
source("R/Enrichment_TG_based.R")
source("R/global_variable.R")
source("R/lib.R")

# load MGI background and pathway gene set
background_MGI <- read.csv("data/raw_data/background/background_MGI.csv")
path_name <- "Reactome" # or "hallmark"
pathway <- read.csv(paste0("data/raw_data/pathway/", path_name, "/", path_name, "_gene.csv"), header = TRUE)


background_MGI <- data.table::as.data.table(background_MGI)
pathway <- data.table::as.data.table(pathway)

# generate TG from a list of miRNAs
##  example
miRs <- c("hsa-miR-210-3p", "hsa-miR-139-3p", "hsa-miR-96-5p", "hsa-miR-195-5p", "hsa-miR-30a-3p", "hsa-miR-28-3p", "hsa-miR-18a-5p", "hsa-miR-143-3p", "hsa-miR-1307-5p", "hsa-miR-17-5p", "hsa-miR-30c-2-3p", "hsa-miR-183-5p", "hsa-miR-141-3p", "hsa-miR-130b-3p", "hsa-miR-139-5p", "hsa-miR-145-3p", "hsa-miR-33a-5p", "hsa-miR-19a-3p", "hsa-miR-30e-3p", "hsa-miR-1-3p", "hsa-miR-671-5p", "hsa-miR-28-5p",   "hsa-miR-93-5p", "hsa-miR-519a-5p", "hsa-miR-20a-5p", "hsa-miR-26a-5p", "hsa-miR-3199",  "hsa-miR-425-5p", "hsa-miR-516a-5p", "hsa-miR-590-5p" )
TG <- background_MGI[miRNA %in% miRs, unique(gene)]

# run TG-ORA
result <- TG_ORA(TG = TG, pathway = pathway, pvalueCutoff = 1)
```
