---
title: "TG-Score"
author: "Zhesi"
date: "2026-09-18"
output: html_document
---

This is the tutorial to implement TG-Score enrichment analysis method.\
Before running, please set up your working directory

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/scratch/project_2011179/code/github/")  # Please change your woking directory here.
```

Note: in .R script, please set your working directory by `setwd("[your directory]")`.

You can run TG-Score enrichment analysis in three ways:\
**A.** run through integrated miREA() function.\
**B.** run by providing input gene list through miRNA differential expression analysis result.\
**C.** run by specifying an interested miRNA list.\
In the following tutorial, we will go through each accordingly.

**Required input:**\
**1. background_MGI:** A two-column data frame with column names miRNA and gene.\
**2. pathway:** A two-column data frame with column names pathway and gene.\
**3. gene_DEdata:** A four-column data frame containing gene, log2FoldChange, stat, and padj.\
**4. mir_DEdata (required in A, B):** A four-column data frame containing miRNA, log2FoldChange, stat, and padj.\
**5. miR (required in C):** a character vector containing miRNAs of interest.

### A. run through miREA().

```{r}
load("function.RData") # it contains all functions included in the R/.
source("R/lib.R")

methods <- "TG_Score"

mir_DEdata <- read.csv("data/raw_data/cancer_data/DEmiR/BLCA_DEmiR.csv")
mir_DEdata <- mir_DEdata %>% select(miRNA, log2FoldChange, stat, padj)
gene_DEdata <- read.csv("data/raw_data/cancer_data/DEG/BLCA_DEG.csv")
gene_DEdata <- gene_DEdata %>% select(gene, log2FoldChange, stat, padj)

path_name <- "Reactome" # or "hallmark"
pathway <- read.csv(paste0("data/raw_data/pathway/", path_name, "/", path_name, "_gene.csv"), header = TRUE)

input_data <- get_all_input_data(methods = methods, pathway = pathway, mir_DEdata = mir_DEdata, gene_DEdata = gene_DEdata)
result <- miREA(methods = methods, input_data = input_data)
```

It will generate a list named result, which contains the enrichment analysis result (`result$result$TG_Score`), calculation time (`result$time`), methods (`result$valid_methods`), and parameters (`result$parameter`).

### B. run by providing miRNA differential expression analysis result.

Load miRNA differential expression analysis result, identify DEmiRs, transform to DEmiRs' target genes and rank the TG list by genes' differential expression analysis statistics.

```{r}
# load functions
source("R/Enrichment_TG_based.R")
source("R/global_variable.R")
source("R/lib.R")

# load MGI background and pathway gene set
background_MGI <- read.csv("data/raw_data/background/background_MGI.csv")
path_name <- "Reactome" # or "hallmark"
pathway <- read.csv(paste0("data/raw_data/pathway/", path_name, "/", path_name, "_gene.csv"), header = TRUE)


background_MGI <- data.table::as.data.table(background_MGI)
pathway <- data.table::as.data.table(pathway)

# load miR and gene DE data
mir_DEdata <- read.csv("data/raw_data/cancer_data/DEmiR/BLCA_DEmiR.csv")
mir_DEdata <- data.table::as.data.table(mir_DEdata)
gene_DEdata <- read.csv("data/raw_data/cancer_data/DEG/BLCA_DEG.csv")
gene_DEdata <- data.table::as.data.table(gene_DEdata)

# build geneList (ranked DEmiRs' target genes list)
DEmiRs <- mir_DEdata[abs(log2FoldChange) > 1 & padj < 0.05, unique(miRNA)]
TG <- background_MGI[miRNA %in% DEmiRs, unique(gene)]
geneList <- gene_DEdata[gene %in% TG,
                    setNames(stat, gene)]
geneList <- sort(geneList, decreasing = TRUE)

# run TG-Score
result <- TG_Score(geneList = geneList, pathway = pathway, pvalueCutoff = 1)
```

### C. run by specifying an interested miRNA list.

if you have a list of interested miRNA directly, you could transform interested miRNAs to their target genes as input gene list.

```{r}
# load functions
source("R/Enrichment_TG_based.R")
source("R/global_variable.R")
source("R/lib.R")

# load MGI background and pathway gene set
background_MGI <- read.csv("data/raw_data/background/background_MGI.csv")
path_name <- "Reactome" # or "hallmark"
pathway <- read.csv(paste0("data/raw_data/pathway/", path_name, "/", path_name, "_gene.csv"), header = TRUE)

background_MGI <- data.table::as.data.table(background_MGI)
pathway <- data.table::as.data.table(pathway)

# load gene DE data
gene_DEdata <- read.csv("data/raw_data/cancer_data/DEG/BLCA_DEG.csv")
gene_DEdata <- data.table::as.data.table(gene_DEdata)

# generate ranked target gene list from a list of miRNAs
##  example
miRs <- c("hsa-miR-210-3p", "hsa-miR-139-3p", "hsa-miR-96-5p", "hsa-miR-195-5p", "hsa-miR-30a-3p", "hsa-miR-28-3p", "hsa-miR-18a-5p", "hsa-miR-143-3p", "hsa-miR-1307-5p", "hsa-miR-17-5p", "hsa-miR-30c-2-3p", "hsa-miR-183-5p", "hsa-miR-141-3p", "hsa-miR-130b-3p", "hsa-miR-139-5p", "hsa-miR-145-3p", "hsa-miR-33a-5p", "hsa-miR-19a-3p", "hsa-miR-30e-3p", "hsa-miR-1-3p", "hsa-miR-671-5p", "hsa-miR-28-5p",   "hsa-miR-93-5p", "hsa-miR-519a-5p", "hsa-miR-20a-5p", "hsa-miR-26a-5p", "hsa-miR-3199",  "hsa-miR-425-5p", "hsa-miR-516a-5p", "hsa-miR-590-5p" )
TG <- background_MGI[miRNA %in% miRs, unique(gene)]
geneList <- gene_DEdata[gene %in% TG,
                    setNames(stat, gene)]
geneList <- sort(geneList, decreasing = TRUE)

# run TG-Score
result <- TG_Score(geneList = geneList, pathway = pathway, pvalueCutoff = 1)
```
